---
title: "Using the NonsmoothPath package"
author: "Qiuyu Gu"
date: "`r Sys.Date()`"
#output: rmarkdown::html_vignette
output: pdf_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `NonsmoothPath` package contains the implementation of the path-following algorithm to solve case weight adjusted problem
\begin{equation}
	\label{eq:case_weight_adjusted_setup}
	(\beta_{0,w}, \beta_w) = \text{argmin}_{\beta_0 \in \mathbb{R}, \beta \in \mathbb{R}^{p}} \sum_{i \neq i^*} f(g_i(\beta_0, \beta)) + w f(g_{i^*}(\beta_0, \beta)) + \frac{\lambda}{2} \|\beta\|_2^2,
\end{equation}
where $g_i(\beta_0, \beta) = a_i \beta_0 + b_i^{\top}\beta + c_i$ for some $a_i, c_i \in \mathbb{R}, b_i \in \mathbb{R}^p$, and $f$ is a piecewise linear function with one knot at 0, i.e. $f(r) = \alpha_0 \max(r, 0) + \alpha_1 \max(-r, 0)$  for some $\alpha_0 \ge 0, \alpha_1 > 0$. The term $f(g_i(\beta_0, \beta))$ can be viewed as the loss function for $i^{th}$ case. If we consider the quantile regression with the parameter $\tau$, then 
\begin{equation*}
	a_i = -1, b_i = -x_i, c_i = y_i, \alpha_0 = \tau, \alpha_1 = 1 - \tau.
\end{equation*}
And if we consider SVM, then
\begin{equation*}
	a_i = y_i, b_i = y_i x_i, c_i = -1, \alpha_0 = 0, \alpha_1 = 1. 
\end{equation*} 
The solution when $w = 1$ corresponds to the full data solution and we assume it is available to us. The algorithm computes the exact solution path of \eqref{eq:case_weight_adjusted_setup}, which is shown to be piecewise linear in $w$, as the parameter $w$ changes from 1 to 0.

## Using `NonsmoothPath` functions
For each given data and specified model, function `nonsmooth_path` in `NonsmoothPath` package computes solution path of \eqref{eq:case_weight_adjusted_setup} and outputs the

1. breakpoints, \

2. $\beta_0, \beta$ and dual variable ($\theta$) at these breakpoints. \

Next, We use quantile regression as an example to show how to apply `nonsmooth_path` function to compute solution path for case weight adjusted problem \eqref{eq:case_weight_adjusted_setup}. First we generate data from the standard linear model
$$
y_i = x_i^{\top} \beta + \epsilon_i,
$$
where covariates $\{x_{ij}: i = 1, \cdots, n, \ j = 1, \cdots, p\}$, coefficients $\{\beta_1, \cdots, \beta_p]\}$ and errors $\{\epsilon_i: i =1, \cdots, n\}$ are independently generated from standard normal distribution. Here tau is the quantile parameter in quantile regression and lam is the regularization parameter for $L_2$ penalty.

```{r}
# n = 100
# p = 200
# lam = 50
# tau = 0.1
# beta_true = rnorm(p)
# X = matrix(0, ncol = p, nrow = n)
# for (i in 1:n)
#   X[i, ] = rnorm(p)
# eps = rnorm(n)
# Y = X %*% beta_true + eps
```

To compute solution path for the generated data under quantile regression, we can call the function `nonsmooth_path`. All we need to do is to specify the `class = "quantile"` and the value for `obs_index` argument, which is the index of the case we attach parameter $w$. 
```{r}
# obs_index = 1
# quantile_path = nonsmooth_path(X, Y, lam, obs_index, class = "quantile", tau = tau)
```

The function outputs a list, which contains all the breakpoints and corresponding $\beta_0, \beta$ and $\theta$. We can access these values through
```{r}
# w_vec = quantile_path$W_vec
# beta_0_vec = quantile_path$Beta_0
# beta_mat = quantile_path$Beta
# theta_mat = quantile_path$Theta
```

To make it work for support vector machine (svm), the following changes are needed: \

1. Set `class = "svm"` \

2. Do not need to specify values for tau. \
